---
title: "Interactive Plotly Presentation"
author: "Enxu Lu"
date: "8/23/2026"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
    transition: 0.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Document Creation Date

**Created on:** 8/23/2026

---

## Introduction

This presentation demonstrates interactive plots created with **Plotly** in R Markdown.

Plotly is a JavaScript library that creates interactive, web-based visualizations.

### Key Features

- Interactive zoom and pan
- Hover tooltips for data details
- Dynamic resizing
- Export to PNG/SVG

---

## Interactive Scatter Plot

```{r plotly-scatter}
library(plotly)
library(ggplot2)

# Create sample data
set.seed(123)
data <- data.frame(
    x = rnorm(100, mean = 50, sd = 15),
    y = rnorm(100, mean = 60, sd = 12),
    category = sample(c("Group A", "Group B", "Group C"), 100, replace = TRUE)
)

# Create interactive scatter plot
plot_ly(
    data = data,
    x = ~x,
    y = ~y,
    color = ~category,
    type = 'scatter',
    mode = 'markers',
    marker = list(size = 12),
    text = ~paste("X:", round(x, 2), "<br>Y:", round(y, 2)),
    hoverinfo = 'text'
) %>%
    layout(
        title = "Interactive Scatter Plot",
        xaxis = list(title = "X Variable"),
        yaxis = list(title = "Y Variable")
    )
```

---

## Interactive 3D Scatter Plot

```{r plotly-3d}
# 3D scatter plot
set.seed(456)
data3d <- data.frame(
    x = rnorm(80, mean = 0, sd = 1),
    y = rnorm(80, mean = 0, sd = 1),
    z = rnorm(80, mean = 0, sd = 1),
    cluster = sample(c("Cluster 1", "Cluster 2", "Cluster 3"), 80, replace = TRUE)
)

plot_ly(
    data = data3d,
    x = ~x,
    y = ~y,
    z = ~z,
    color = ~cluster,
    type = 'scatter3d',
    mode = 'markers',
    marker = list(size = 8)
) %>%
    layout(
        title = "3D Scatter Plot",
        scene = list(
            xaxis = list(title = "X"),
            yaxis = list(title = "Y"),
            zaxis = list(title = "Z")
        )
    )
```

---

## Interactive Line Chart

```{r plotly-line}
# Time series data
time_data <- data.frame(
    date = seq.Date(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"),
    sales = c(120, 135, 140, 155, 160, 175, 190, 185, 200, 210, 195, 220),
    expenses = c(80, 85, 90, 95, 100, 110, 115, 120, 125, 130, 135, 140)
)

plot_ly() %>%
    add_trace(
        data = time_data,
        x = ~date,
        y = ~sales,
        name = 'Sales',
        type = 'scatter',
        mode = 'lines+markers',
        line = list(width = 3),
        marker = list(size = 8)
    ) %>%
    add_trace(
        data = time_data,
        x = ~date,
        y = ~expenses,
        name = 'Expenses',
        type = 'scatter',
        mode = 'lines+markers',
        line = list(width = 3, dash = 'dash'),
        marker = list(size = 8)
    ) %>%
    layout(
        title = "Monthly Sales and Expenses",
        xaxis = list(title = "Date"),
        yaxis = list(title = "Amount ($)")
    )
```

---

## Interactive Bar Chart

```{r plotly-bar}
# Bar chart data
bar_data <- data.frame(
    city = c("New York", "London", "Tokyo", "Paris", "Sydney", "Berlin"),
    population = c(8419000, 8982000, 13960000, 11060000, 5312000, 3645000)
)

plot_ly(
    data = bar_data,
    x = ~city,
    y = ~population,
    type = 'bar',
    color = ~city,
    text = ~paste("Population:", format(population, big.mark = ",")),
    hoverinfo = 'text'
) %>%
    layout(
        title = "City Populations",
        xaxis = list(title = "City"),
        yaxis = list(title = "Population")
    )
```

---

## Interactive Histogram

```{r plotly-histogram}
# Histogram with normal curve overlay
hist_data <- rnorm(1000, mean = 70, sd = 10)

plot_ly(
    x = ~hist_data,
    type = 'histogram',
    nbinsx = 30,
    marker = list(
        color = 'rgba(100, 200, 255, 0.6)',
        line = list(color = 'rgba(0, 0, 0, 0.3)', width = 1)
    )
) %>%
    layout(
        title = "Distribution of Values",
        xaxis = list(title = "Value"),
        yaxis = list(title = "Frequency")
    )
```

---

## Interactive Heatmap

```{r plotly-heatmap}
# Correlation matrix heatmap
set.seed(789)
corr_data <- matrix(runif(25, -1, 1), nrow = 5)
diag(corr_data) <- 1
colnames(corr_data) <- rownames(corr_data) <- c("A", "B", "C", "D", "E")

plot_ly(
    z = corr_data,
    x = colnames(corr_data),
    y = rownames(corr_data),
    type = 'heatmap',
    colorscale = 'RdBu',
    reversescale = TRUE,
    hovertemplate = 'x: %{x}<br>y: %{y}<br>correlation: %{z:.2f}<extra></extra>'
) %>%
    layout(
        title = "Correlation Matrix",
        xaxis = list(title = ""),
        yaxis = list(title = "")
    )
```

---

## Interactive Box Plot

```{r plotly-boxplot}
# Boxplot by category
set.seed(101)
box_data <- data.frame(
    value = c(rnorm(50, 100, 15), rnorm(50, 120, 20), rnorm(50, 90, 10)),
    group = rep(c("Group 1", "Group 2", "Group 3"), each = 50)
)

plot_ly(
    data = box_data,
    x = ~group,
    y = ~value,
    type = 'box',
    color = ~group,
    boxpoints = 'all',
    jitter = 0.3,
    pointpos = -1.8
) %>%
    layout(
        title = "Box Plot by Group",
        xaxis = list(title = "Group"),
        yaxis = list(title = "Value")
    )
```

---

## Interactive Pie Chart

```{r plotly-pie}
# Pie chart
pie_data <- data.frame(
    category = c("Category A", "Category B", "Category C", "Category D"),
    value = c(35, 25, 20, 20)
)

plot_ly(
    data = pie_data,
    labels = ~category,
    values = ~value,
    type = 'pie',
    textinfo = 'label+percent',
    hoverinfo = 'label+value',
    marker = list(
        colors = c('rgba(255, 99, 132, 0.8)',
                   'rgba(54, 162, 235, 0.8)',
                   'rgba(255, 206, 86, 0.8)',
                   'rgba(75, 192, 192, 0.8)')
    )
) %>%
    layout(
        title = "Category Distribution"
    )
```

---

## Interactive Density Plot

```{r plotly-density}
# Density plot with multiple groups
set.seed(202)
density_data <- data.frame(
    value = c(rnorm(200, 50, 10), rnorm(200, 70, 12), rnorm(200, 60, 8)),
    group = rep(c("Group A", "Group B", "Group C"), each = 200)
)

plot_ly(
    data = density_data,
    x = ~value,
    color = ~group,
    type = 'histogram',
    histnorm = 'probability density',
    nbinsx = 30,
    opacity = 0.6
) %>%
    layout(
        title = "Density Plot by Group",
        xaxis = list(title = "Value"),
        yaxis = list(title = "Density"),
        barmode = 'overlay'
    )
```

---

## Interactive Map with Plotly

```{r plotly-map}
# Map with Plotly (scattergeo)
map_data <- data.frame(
    city = c("New York", "London", "Tokyo", "Sydney", "Paris", "Moscow"),
    lat = c(40.712, 51.507, 35.689, -33.868, 48.857, 55.755),
    lon = c(-74.006, -0.127, 139.692, 151.209, 2.352, 37.618),
    population = c(8419000, 8982000, 13960000, 5312000, 11060000, 11920000)
)

plot_ly(
    data = map_data,
    lat = ~lat,
    lon = ~lon,
    type = 'scattergeo',
    mode = 'markers',
    marker = list(
        size = ~sqrt(population / 1000000) * 5,
        color = 'red',
        opacity = 0.8
    ),
    text = ~paste(
        city, "<br>",
        "Population:", format(population, big.mark = ",")
    ),
    hoverinfo = 'text'
) %>%
    layout(
        title = "Major Cities Map",
        geo = list(
            projection = list(type = 'orthographic'),
            showland = TRUE,
            landcolor = 'rgb(200, 200, 200)',
            countrycolor = 'rgb(150, 150, 150)'
        )
    )
```

---

## Summary

### Plotly Features Used

| Plot Type | Function | Interaction |
|-----------|----------|-------------|
| Scatter | `plot_ly(type = 'scatter')` | Zoom, hover, drag |
| 3D Scatter | `plot_ly(type = 'scatter3d')` | Rotate, zoom, hover |
| Line | `plot_ly(type = 'scatter', mode = 'lines')` | Hover, zoom |
| Bar | `plot_ly(type = 'bar')` | Hover |
| Histogram | `plot_ly(type = 'histogram')` | Hover |
| Heatmap | `plot_ly(type = 'heatmap')` | Hover |
| Boxplot | `plot_ly(type = 'box')` | Hover, zoom |
| Pie | `plot_ly(type = 'pie')` | Hover |
| Density | `plot_ly(type = 'histogram', histnorm = 'probability density')` | Hover |
| Map | `plot_ly(type = 'scattergeo')` | Zoom, rotate, hover |

---

## About This Document

| Item | Information |
|------|-------------|
| **Created on** | 8/23/2026 |
| **Author** | Enxu Lu |
| **Purpose** | Peer-graded assignment for Coursera |
| **Tools Used** | R Markdown, Plotly, GitHub Pages |
| **R Packages** | plotly, ggplot2 |

---

## References

1. Plotly for R: https://plotly.com/r/
2. R Markdown: https://rmarkdown.rstudio.com/
3. GitHub Pages: https://pages.github.com/